Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-09 paper.tex
a bit more content in experiments section

\usepackage[bookmarks=false]{hyperref}

@@

\subsubsection{Applications} % 1 page

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline that consisted exclusively of map 
stages. To would allow us to create experiments to tweak data locality, 
in-memory computing and lazy evaluation easily. Incrementation was selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

For our first and second experiment\todo{better split into two paragraphs}, we wanted to determine the effects of in-memory 
processing on increasing data size and task duration. In other words, the 125 
chunk BigBrain 
image was increment 1, 10 and 100 times, such that the total processed data 
size increased from 75GB, at 1 iteration, to 7500GB at 100 iterations. In order
to validate the model, the D/C was varied to fall in different regions of 
Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
and 1.9MB/s\todo{present this as a table?}. 80 and 1.9MB/s satisfied the 
inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
Lustre. These D/C ratios were applied to each iteration experiment, such that 
we were able to get both the effect of increasing the total amount of data, in 
addition to, increasing the task duration for a given data size.

@@

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed at 125. Similarly to the previous experiment, the total 
sequential compute time was fixed, however, due to varying size in total data 
processed, the D/C ratio varied. Once again, we ensure that the D/C ratio fell
in multiple different regions of the inequality.

>>>>

\usepackage[bookmarks=false]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

\subsubsection{Applications} % 1 page
\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require $x$, a sleep delay in float
    \Require $fs$, filesystem to save image to (i.e. mem, tmpfs, local disk, lustre). 
    \ForEach{$i \in iterations$}
    \ForEach{$chunk \in image$}
        \State $chunk\gets chunk+1$
        \State sleep $x$
        \State save $chunk$ to $fs$
    \EndFor
    \EndFor
\end{algorithmic}
\end{algorithm}

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

For our first incrementation experiment, we investigated the effects of
in-memory computing on varying total data size. to do this, we increased the 
number of incrementation iterations from 1, 10 and 100 times. The total data 
size would then increased from 75GB, at 1 iteration, to 7500GB, at 100 
iterations. In order
to validate the model, the D/C was varied to fall in different regions of 
Equation~\ref{eq:page-cache-inequality}. The D/C selected were 256, 178.6, 80 
and 1.9MB/s\todo{present this as a table?}. 80 and 1.9MB/s satisfied the 
inequality for local disk, whereas only 1.9MB/s satisfied the inequality for 
Lustre. These D/C ratios were applied to each iteration experiment, such that 
we were able to get both the effect of increasing the total amount of data, in 
addition to, increasing the task duration for a given data size.

We combined the first experiment with our second experiment such the we could 
evaluate the effects of in-memory computing on varying task duration. If the 
page cache has sufficient time to flush, it should be expected that using 
in-memory computing should have no increased speed-up over disk. We maintained 
the D/C ratios as those mentioned above, but rather than applying only one to
a different number of iterations, we applied all four D/C to all numbers of 
iterations. That is 1 iteration, 10 iterations and 100 iterations were all 
processed with using the aforementioned D/C ratios.

@@

For our fourth and final incrementation experiment, we investigated the effects 
of the strategies on different image sizes. We selected the 75GB BigBrain, the
38GB half BigBrain and the 13M T1W MRI image for this experiment. The number of
chunks was fixed at 125. Similarly to the previous experiment, the total 
sequential compute time was fixed, however, due to varying size in total data 
processed, the D/C ratio varied. Once again, we ensured that the D/C ratio fell
in multiple different regions of the inequality. The D/C ratio ranged from 
348.8 MB/s for BigBrain and 174.4MB/s for half Bigbrain, to 0.1MB/s for the MRI 
image. Only the 0.1MB/s MRI satisfied the inequality for both Lustre and local 
disk.

Since data-locality is not normally preserved in Nipype (a new SLURM allocation 
is requested for each processed task), we instrumented Nipype to ensure data 
locality. That is, for each chunk partition, we requested a SLURM allocation to
process the entire pipeline in parallel, using Nipype's MultiProc scheduler, on 
a given node. This was possible as no communication was required between the 
processed chunks.
